---
title: "IBD-cohort"
author: "Yilu"
date: "06/10/2021"
output: html_document
---

# 1.load packages
```{r,echo=FALSE}
library(CEMiTool)
library(dplyr)
library(janitor)
library(doParallel) ##
registerDoParallel(cores=4)
library(preprocessCore)
```

# 2.laod data
```{r, echo=FALSE}
IBD_meta<-read.table('/Users/yihuawang/Downloads/IBD_sub.txt',header = T,sep = '\t') %>% 
  filter(Diagnosis=='CD' | Diagnosis=='UC')%>%
  rename('SampleName'='IID',
         'Class'='Diagnosis')%>%
  arrange(., Class)

print(table(IBD_meta$Class))
#CD  UC 
#681 368

IBD<-read.table('/Users/yihuawang/Downloads/GENEPY_JULY2021_all.matrix.txt',header = T,sep = '\t',stringsAsFactors = F) %>%
  t() %>% 
  as.data.frame() %>%
  row_to_names(row_number = 1)%>%
  select(IBD_meta$SampleName)
IBD[,1:1049] <- sapply(IBD[,1:1049], as.numeric) # 9130 genes with 0 expression in IBD[,1]
```

# 3.check basic feature of this matrix
```{r}
#### PCA ####
library(factoextra)
IBD_filter<-IBD[rowSums(IBD > 2),] # remove rows with low values
data_t <- t(as.matrix(IBD_filter))
variableL <- ncol(data_t)
pca <- prcomp(data_t[,1:variableL], scale=T)
#fviz_pca_ind(pca, repel=T)  
data_t<-as.data.frame(data_t)
data_t$conditions<-as.character(IBD_meta$Class)
#pdf('PCA_withoutremovebatch.pdf',width = 8,height = 8)
fviz_pca_ind(pca, col.ind=data_t$conditions, mean.point=F, addEllipses = T, legend.title="Dataset",
             geom.ind = "point")+scale_shape_manual(values=seq(0,17))+theme_classic()

```

# 4.Sample correlation heatmap
```{r}
res <- cor(IBD_filter)
col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(x = res, col = col, symm = TRUE)

IBD2<-IBD
colnames(IBD2)<-IBD_meta$Class
res <- cor(IBD2)
col<- colorRampPalette(c("blue", "white", "red"))(20)
heatmap(x = res, col = col, symm = TRUE)
pdf('cor.pdf',width = 20,height = 20)
heatmap(x = res, col = col, symm = TRUE)
dev.off()
```

# 5. Load GMT and PPI files
```{r}
mads=apply(IBD,1,mad) ### measured by median absolute deviation
IBD_mad=IBD[rev(order(mads))[1:5000],] ### top 5000 variable genes
#IBD_mad = sweep(IBD_mad,1, apply(IBD_mad,1,median,na.rm=T))
#write(rownames(IBD_mad),'IBD_mad_varing_genes.txt',sep = '\t')
# read Interactions file
int_df<-read.csv('/Users/yihuawang/Downloads/String Network - 1 default edge copy.csv',header = T,sep = ',')
head(int_df)

# read GMT file
gmt_hallmark <- read_gmt('/Users/yihuawang/Downloads/h.all.v7.4.symbols.gmt')
gmt_kegg <- read_gmt('/Users/yihuawang/Downloads/c2.cp.kegg.v7.4.symbols.gmt')
gmt_go <- read_gmt('/Users/yihuawang/Downloads/c5.go.v7.4.symbols.gmt')
```

# 6.Failed to run weighted correlation network analysis: No modules found
```{r}
cem <- cemitool(as.data.frame(IBD_mad),IBD_meta,gmt_go,interactions=int_df,
                set_beta=2,filter=TRUE, plot=F, verbose=TRUE)


# create report as html document
generate_report(cem, directory="./Report",force=TRUE)

# write analysis results into files
write_files(cem, directory="./Tables",force=TRUE)

# save all plots
save_plots(cem, "all", directory="./Plots",force=TRUE)
```
# cemitool tips
1. Check if your data contain at least 20 samples, which we consider to be the very least for any meaningful analysis.| pass
2. Make sure your data were normalized adequately prior to running CEMiTool. Normalization techniques which have been used successfully in the past include RPKM, FPKM, TMM, quantile... | need to consider
3. You can see if your data have too much dependence between the mean and variance. If this is the case, using the apply_vst = TRUE parameter can help. | need to consider
4. Another possibility is that CEMiTool was too stringent in the beta selection step of analysis. You can check this by using plot_beta_r2(cem), and then show_plot(cem, "beta_r2"). You can then manually select a beta value with the set_beta value.
5. Finally, you can make CEMiTool select a beta value based on the number of samples you have with the force_beta parameter.
6. Please see the WGCNA FAQ for more info and caveats on the last point, specifically item 6.

# 7.fold change values of top5000 varying genes 
```{r}
IBD_mad_CD<-IBD_mad[,IBD_meta[IBD_meta$Class== 'CD',1]]
IBD_mad_UC<-IBD_mad[,IBD_meta[IBD_meta$Class== 'UC',1]]
IBD_mad_FC<-as.data.frame(IBD_mad_CD[,1])
colnames(IBD_mad_FC)<-'FC'
rownames(IBD_mad_FC)<-rownames(IBD_mad_CD)
IBD_mad_FC$FC<-log2(rowMeans(IBD_mad_CD)/rowMeans(IBD_mad_UC))
```

# 8.Create Gephi input files
```{r}
IBD_mad_FC$Gene<-rownames(IBD_mad_FC)

int_df<-read.csv('/Users/yihuawang/Downloads/String Network - 1 default edge copy.csv',header = T,sep = ',')

input_node<-as.data.frame(unique(c(int_df$gene1symbol,int_df$gene2symbol)))
colnames(input_node)<-'Gene'
input_node<-merge(input_node,IBD_mad_FC,by='Gene')
write.table(input_node,'input_node.csv',sep = ',')
```